---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Notes to contributors

Stick to the structure, design choices and style conventions described below.
For questions: please contact the [author](mailto:manuel.steiner@uni-wuerzburg.de).

### Structure

The package structure is best understood with reference to a (hierarchically organized) company. In analogy to company departments, two separate "function departments" 
exist:

- **Estimation functions**
- **Postestimation functions**

Each department is hierarchically structured. All functions that do not belong
to one of the departments are considered **utility functions**. To stay in the
picture, they are best understood as external consultants in charge of one
specific task that is not meaningfully classified as belonging to one of the two
departments.

#### Estimation functions:

Estimation functions are all functions involved in the estimation/calculation 
of the quantities of the measurement and the path model. There are:

- **2 toplevel functions** (`csem()` and `cca()`). These functions + the
  postestimation function `summary()` (see below) should be sufficient for the end
  user 95% of the time. Both functions eventually call `foreman()`.
- **1 midlevel function**, `foreman()`, that acts like a foreman by collecting all
  (estimation) tasks, distributing them to lower level functions, and 
  eventually recollecting all of their results.
- **Lowlevel (helper) functions** that perform one specific task. A distinction
  is made between:
    - **Exported helper functions**: Functions that are exported to be applied
      directly by the end user if needed (e.g. `parseModel()` or
      `calculateWeightsPLS()`).
    - **Internal helper functions**: Functions that are not exported. These functions 
      can be accessed via `csem::` but they are not generally useful to the end user
      (e.g. `calculateConstructVCV()` or `classifyConstructs()`). 
      
#### Postestimation functions
**Postestimation functions**. Postestimation functions are generic functions
to be applied to an object resulting from a call to `csem()` or `cca()`. Depending
on the class of the resulting object (e.g. `cSEMResults` or `cSEMSummary`)
different methods are dispatched. Most postestimation functions like `test()`
(not yet implemented) act like a foreman by the collect-distribute-recollect 
strategy as implemented in the `foreman()` function. 
Therefore postestimation functions are generally also hierarchical. 
Currently only a subset of these functions is implemented:

- 1 Top-level function: 
  - `summary()` with class `cSEMSummary`
- Mid-level functions:
  - `status()` with class `cSEMStatus`
  - `fit()` with class `cSEMFit`
  - `coef()` with class `cSEMCoef`
  - `effects()` with class `cSEMEffects`
  - `test()` with class `cSEMTest`
- Low-level functions:
  - `test_fit()` with class `cSEMTestFit`
  - `test_MICOM()` with class `cSEMTestMICOM`
  - `test_MGA()` with class `cSEMTestMGA`

Each function has (will eventually have) a distinct class and a corresponding 
`print` method.

### Design choices

- The only OO system used is the S3 system. No S4 classes will be allowed! (Strong
  recommendation by Yves Rosseel)
- Whenever you subset a matrix using `[` use: `[..., ..., drop = FALSE]` to avoid
  accidentally droping the `dim` attributes.
  
### Style/Naming

Stick to [this styleguide](http://style.tidyverse.org/) with the following
exceptions/additions:

1. Function and class names are always CamelCase. Function names should contain 
   a verb followed by a noun like: `processData()`, `calculateValue()`.
1. Verbs in function names should be consistent across the whole package. Avoid to
   mix synonyms. Example: `computeValue()` vs. `calculateValue()`.
   This package always uses `calculate` instead of `compute`. Similarly, `method` 
   vs e.g. `approach`. This package always uses `approach` instead of `method`.
1. Use plural in function/object names if the main output is more than one element, 
   like `scaleWeights()`, `calculateProxies()`, `handleArgs()` etc. but stick to
   singular in other cases like `parseModel()`.
1. Strive for meaningful argument names even if they are a bit longer than usual.
   People are much better at remembering arguments like `respect_structural_model`
   compared to something like `resp_sm`. Naming should also be consistent if possible.
   For example: any argument that describes a method or an approach should be named
   `.appraoch_*`.
1. Argument names always start with a dot to distinguish them from other objects.
1. Indentation: It is OK to align function arguments indented by two spaces below 
   the function name instead of where the function starts if this help with clarity.

```{r eval=FALSE}
## Both ok but second is prefered in this case
calculateInnerWeightsPLS <- function(.S                           = NULL,
                                     .W                           = NULL,
                                     .csem_model                  = NULL,
                                     .PLS_weight_scheme_inner     = c("centroid",
                                                                      "factorial", 
                                                                      "path"),
                                     .PLS_ignore_structural_model = NULL
                                     ) { }

calculateInnerWeightsPLS <- function(
  .S                            = NULL,
  .W                            = NULL,
  .csem_model                   = NULL,
  .PLS_weight_scheme_inner      = c("centroid","factorial", "path"),
  .PLS_ignore_structural_model  = NULL
  ) { }
```

### Matrices

- Whenever possible: variables belong in columns, observations belong to rows.
  This implies: whenever a matrix contains observations, variables are in columns, 
  no matter if they are indicators, proxies, errors or anything else.
- Covariance matrices: indicators **always** belong to columns and proxies to 
  rows, i.e., the matrix of weights $W$ for PLS is therefore $(J \times K)$ 
  where $J$ is the number of proxies and $K$ the number of indicators.
- Naming: matrices within the package should be named according to the naming schemes 
  of the related SEM literature. 

### Helper functions

Exported helper functions should be written as autonomous as possible in a sense 
that they can be used without having to jump to a mother function in order to allow researchers using the package to use helper function the way the need it. 
Flexibility will come at the price of code repetition (i.e. most exported helper 
functions will have to have a `parseModel()` + `processDate()` statement at the 
beginning) to make them autonomous.

Internal helper functions on the other hand do not need to be autonomous. Here
code compactness is preferred.

### Arguments

All arguments used by any of the functions in the package (including internal
functions) are centrally collected in the file `zz_arguments.R`.
Whenever a new argument is introduced:

1. Add the new argument + a description to the `cSEMArguments` list by writing
  `@param <argument> "Some description". "Default to xxx".`
1. Add the argument to the `args`-list in `args_default()` function and provide a
   default value. 
1. Add the argument to the function you want to use it in. Order arguments
   according to their importance (i.e. `.data` and `.model` always come first).
   if there is one otherwise use alphabetical order.
