---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# cSEM: a package for composite-based SEM <img src='man/figures/cSEMsticker.svg' align="right" height="200" /></a>

[![CRAN status](https://www.r-pkg.org/badges/version/cSEM)](https://cran.r-project.org/package=cSEM)
[![Build Status](https://travis-ci.com/M-E-Rademaker/cSEM.svg?branch=master)](https://travis-ci.com/M-E-Rademaker/cSEM)

WARNING: THIS IS WORK IN PROGRESS. BREAKING CHANGES TO THE API ARE VERY LIKELY. 
         Do not use the package before the first stable relase (which will
         be 0.0.1, most likely in mid 2019).

## Purpose

Estimate, analyse, test, and study linear, nonlinear, hierachical and
multigroup structural equation models using composite-based approaches and procedures including estimation 
techniques such as partial least squares path modelling (PLS) and its derivatives
(PLSc, ordPLSc, robustPLSc), generalized structured component analysis (GSCA), generalized structured component 
analysis with uniqueness terms (GSCAm), generalized canonical correlation analysis (GCCA),
unit weights (sum score) and fixed weights, as well as several tests and typical postestimation 
procedures (e.g., assess the model fit, compute direct, indirect and total effects).

## Installation

```{r, eval = FALSE}
# Currently only a development version from GitHub is available:
# install.packages("devtools")
devtools::install_github("M-E-Rademaker/cSEM")
```

## Philopsophy

- User-centred!!
- Easy to use by non-R experts: 
    - One central function `csem()` provides default choices for most of its arguments 
      (similarity to the `sem()` and `cfa()` functions of the [lavaan](http://lavaan.ugent.be/) 
      package is intended).
    - (Eventually...) well documented (Vignettes, HTML output, 
      a website, intro course(s), cheatsheets)
    - Structured output/results  that aims to be "easy"" in a sense that it is
        - ... descriptive/verbose
        - ... easy to export to other environments such as MS Word, Latex files etc. (exportability)
        - ... easy to migrate from/to/between other PLS/VB/CB-based systems (lavaan, semPLS, ADANCO, SmartPLS)
- The package is designed to be flexible/modular enough so that researchers 
  developing new methods can take specific function provided by the package and 
  alter them according to their need without working their way through a chain
  of other functions (naturally this will not always be possible). Modularity is
  largly inspired by the `matrixpls` package.
- Modern in a sense that the package integrates modern developments within 
  the R community. This mainly includes ideas/recommendations/design choices that
  fead into the packages of the [tidyverse](https://github.com/tidyverse/tidyverse).
      
## Basic usage

Models are defined using [lavaan syntax](http://lavaan.ugent.be/tutorial/syntax1.html) with some slight modification.
```{r eval=FALSE}
require(cSEM)
data(satisfaction)
    
## Note: the opeartor "<~" tells cSEM that the construct to its left is modelled
##       as a composite.
##       the operator "=~" tells cSEM that the construct to its left is modelled
##       as a common factor.
    
model <- "
# Structural model
EXPE ~ IMAG
QUAL ~ EXPE
VAL  ~ EXPE + QUAL
SAT  ~ IMAG + EXPE + QUAL + VAL 
LOY  ~ IMAG + SAT

# Measurement model

IMAG <~ imag1 + imag2 + imag3
EXPE <~ expe1 + expe2 + expe3 
QUAL <~ qual1 + qual2 + qual3 + qual4 + qual5
VAL  <~ val1  + val2  + val3
SAT  =~ sat1  + sat2  + sat3  + sat4
LOY  =~ loy1  + loy2  + loy3  + loy4
"
```
Estimation is done using the `csem()` function.
```{r eval=FALSE}
# Estimate using defaults
a <- csem(.data = satisfaction, .model = model)

# This is equal to
csem(
   .data                        = NULL,
   .model                       = NULL,
   .approach_cor_robust         = "none",
   .approach_nl                 = "sequential",
   .approach_paths              = "OLS",
   .approach_weights            = "PLS-PM",
   .conv_criterion              = "diff_absolute",
   .disattenuate                = TRUE,
   .dominant_indicators         = NULL,
   .estimate_structural         = TRUE,
   .id                          = NULL,
   .iter_max                    = 100,
   .normality                   = TRUE,
   .PLS_approach_cf             = "dist_squared_euclid",
   .PLS_ignore_structural_model = FALSE,
   .PLS_modes                   = NULL,
   .PLS_weight_scheme_inner     = "path",
   .reliabilities               = NULL,
   .tolerance                   = 1e-05,
   .resample_method             = "none", 
   .resample_method2            = "none",
   .R                           = 499,
   .R2                          = 199,
   .handle_inadmissibles        = "drop"
   .user_funs                   = NULL,
   .eval_plan                   = "sequential",
   .seed                        = NULL
    )
```

The result is always an object of class `cSEMResults`. Technically, the resulting
object has an additional class attribute (namely `cSEMResults_default`, `cSEMResults_multi`
or `cSEMResults_2ndorder`), however, users usually dont need to worry since postestimation
functions (will eventually) automatically work on all classes. 

```{r eval=FALSE}
## Access elements using `$`. E.g.:
a$Estimates$Loading_estimates 
a$Information$Model
    
## Examine the structure:
listviewer::jsonedit(a, mode = "view") # requires the listviewer package.
    
## Get a summary
summarize(a) 
    
# Alter the model to obtain a linear model:
model <- "
    # Structural model
    EXPE ~ IMAG
    QUAL ~ EXPE
    VAL  ~ EXPE + QUAL
    SAT  ~ IMAG + EXPE + QUAL + VAL
    LOY  ~ IMAG + SAT
    
    # Measurement model
    
    IMAG <~ imag1 + imag2 + imag3
    EXPE <~ expe1 + expe2 + expe3 
    QUAL <~ qual1 + qual2 + qual3 + qual4 + qual5
    VAL  <~ val1  + val2  + val3
    SAT  =~ sat1  + sat2  + sat3  + sat4
    LOY  =~ loy1  + loy2  + loy3  + loy4
    "
    
a <- csem(.data = satisfaction, .model = model)
    
## Apply postestimation functions, e.g.
verify(a) 
    
## Test overall model fit
testOMF(a) # takes roughly 30 seconds

```

Inference can be done by first setting `.resample_method` to 
`"jackkinfe"` or `"bootstrap"` and subsequently using `summarize()` or `infer()`.

```{r eval=FALSE}
b <- csem(.data = satisfaction, .model = model, .resample_method = "bootstrap")
summarize(b)

## Several confidence intervals are implemented, see ?infer():

infer(b, .quantity = c("CI_standard_z", "CI_percentile"))
```

