---
title: "IGSCA Benchmarking"
author: ""
date: today
date-format: long
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 3
    toc-expand: true
    toc-location: left
    code-fold: show
    code-tools:
      source: true
      toggle: true
    code-download: true
    embed-resources: true
    max-width: 1600px
  typst:
    toc: true
bibliography: grateful-refs.bib
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: true

library(RhpcBLASctl)
RhpcBLASctl::blas_set_num_threads(2)
library(profvis)
library(proftools)
library(bench)
library(grateful)
library(sessioninfo)
library(rlang)
rlang::global_handle()
library(lobstr)
cli::pretty_print_code()
library(htmlwidgets)
library(graph)
library(Rgraphviz)
```

# LLM Disclosure

The following describes the use of large language models (LLMs) in the preparation of this document:

- LLMs were used to assist with code drafting, debugging, and editing prose.
- All LLM-generated content was reviewed and verified by the author(s).
- Final responsibility for the accuracy and integrity of this work rests with the author(s).

# Profile Visualization

Here, I use `profvis` to see what might be the bottlenecks in the current git version of the code. 

## Repository Info

```{r}
#| label: git-info
git_hash <- system("git rev-parse HEAD", intern = TRUE)
git_short <- system("git rev-parse --short HEAD", intern = TRUE)
git_branch <- system("git rev-parse --abbrev-ref HEAD", intern = TRUE)
git_msg <- system("git log -1 --format=%s", intern = TRUE)
```

This document was compiled from commit **`r git_short`** on branch **`r git_branch`** (`r git_msg`).
Full SHA: `r git_hash`

```{r}
#| label: profvis-dir
profvis_dir <- here::here(
  "dev/igsca/benchmarking_R",
  paste0(git_branch, "_", format(Sys.Date(), "%Y-%m-%d"))
)
dir.create(profvis_dir, recursive = TRUE, showWarnings = FALSE)
```

## Profile Visualization

```{r}
source(here::here("dev/igsca/benchmarking_R/igsca_benchmark.R"))
```

### Big N Small p

#### proftools

```{r}
pt_bigN_smallp <- proftools::profileExpr(bigN_smallp())
```

```{r,  eval = FALSE}
funSummary(pt_bigN_smallp) |> View()
```

```{r, eval  = FALSE}
callSummary(pt_bigN_smallp) |> View()
```

```{r,  eval = FALSE}
srcSummary(pt_bigN_smallp)
```

```{r,  eval = FALSE}
hotPaths(pt_bigN_smallp)
```

```{r}
plotProfileCallGraph(pt_bigN_smallp)
```

```{r}
flameGraph(pt_bigN_smallp)
```

```{r}
calleeTreeMap(pt_bigN_smallp)
```

#### Profvis 

```{r}
#| output: false
pv_bigN_smallp <- profvis(
  {
    bigN_smallp()
  },
  prof_output = file.path(profvis_dir, "pv_bigN_smallp.Rprof"),
  width = "100%"
)
```

```{r}
pv_bigN_smallp
```


### Small N Big p
#### proftools

```{r}
pt_smallN_smallp <- proftools::profileExpr(smallN_smallp())
```

#### Profvis 

```{r}
#| output: false
pv_smallN_smallp <- profvis(
  {
    smallN_smallp()
  },
  prof_output = file.path(profvis_dir, "pv_smallN_smallp.Rprof"),
  width = "100%"
)
```

```{r}
pv_smallN_smallp
```




### Small N Small p
#### proftools

```{r}
pt_smallN_bigp <- proftools::profileExpr(smallN_bigp())
```

#### Profvis 

```{r}
#| output: false
pv_smallN_bigp <- profvis(
  {
    smallN_bigp()
  },
  prof_output = file.path(profvis_dir, "pv_smallN_bigp.Rprof"),
  width = "100%"
)
```

```{r}
pv_smallN_bigp
```




### Big N Big p
#### proftools

```{r}
pt_bigN_bigp <- proftools::profileExpr(bigN_bigp())
```

```{r}
plotProfileCallGraph(pt_bigN_bigp)
```

```{r}
flameGraph(pt_bigN_bigp)
```

```{r}
calleeTreeMap(pt_bigN_bigp)
```

#### Profvis 

```{r}
#| output: false
pv_bigN_bigp <- profvis(
  {
    bigN_bigp()
  },
  prof_output = file.path(profvis_dir, "pv_bigN_bigp.Rprof"),
  width = "100%"
)
```

```{r}
pv_bigN_bigp
```

### Save profvis outputs

```{r}
#| label: save-profvis
profvis_outputs <- list(
  pv_bigN_smallp  = pv_bigN_smallp,
  pv_smallN_smallp = pv_smallN_smallp,
  pv_smallN_bigp  = pv_smallN_bigp,
  pv_bigN_bigp    = pv_bigN_bigp
)

for (nm in names(profvis_outputs)) {
  htmlwidgets::saveWidget(
    profvis_outputs[[nm]],
    file = file.path(profvis_dir, paste0(nm, ".html")),
    selfcontained = TRUE
  )
}
```

#  List of potential fixes

- Differentiate argument name from parameter itself, potentially triggering look-ups that might slow things down.


- Based on  both the profvis and the interpretation of profvis using Claude, the following functions should be minimized 

```{r, eval = FALSE}
%*% # A lot of time
svd() # Some of the most time
qr() # Some of the most time

ginv() # some time
solve() # some time

norm() # very little time
t()
```

## Potential Action Plan

1. Re-organize functions used for updating weights -- only two. One for updating weights of composites and another for updating weights of common factors. Also fix the assignment of the construct scores as that is computationally costly... [R]
2. Re-organize `updateCBDU()` [R]
3. Make the update of unique scores and unique loadings into a single common function across igsca and GSCA-M [R]
4. Minimize the number of transposes `t()`, `%*%` and `diag` if possible. If there's a large number of matrix multiplications to do in a zone, then consider putting  it all into a single C++ program. Also see if  the use of `MASS::ginv()` can be reduced, as it uses `svd()` and is also computationally expensive. Also, try avoiding memory intensive things where we are writing to a subset of an object. [R]
5. Improve numerical stability of updating of  loadings/path-coefficients [R]
6. Improve numerical stability of updating of weights if possible [R], or else write into C++ [C++]
7. Write update of unique/scores unique-loadings in C++ [C++]
8. Improve numerical stability for how unique scores/unique-loadings is done
9. Reduce the number of random assignments + temporary variables, if possible

# Benchmarking

As long as I'm not comparing between installed vs locally-loaded code packages, this should be legitimate.

Here, I benchmark different approaches for handling specific  bottlenecks. It takes a long time to run, but on my home PC, it seems like 2 threads is optimal, but 1 thread is very close. On the HPC, they disable hyperthreading anyways...


```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(24)
system.time({bigN_bigp()})
```

```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(4)
system.time({bigN_bigp()})
```

```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(2)
system.time({bigN_bigp()})
```

```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(1)
system.time({bigN_bigp()})
```

# Computational Environment


```{r}
#| label: session-info
sessioninfo::session_info(info = "all")
```

## Packages Used

```{r}
#| label: pkg-citations
grateful::cite_packages(output = "paragraph", out.dir = ".")
```


