---
title: "IGSCA Benchmarking"
author: ""
date: today
date-format: long
format:
  html:
    theme:
      dark: darkly
      light: flatly
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools:
      source: true
      toggle: true
    code-download: true
    embed-resources: true
  typst:
    toc: true
bibliography: grateful-refs.bib
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false
library(profvis)
library(proftools)
library(bench)
library(grateful)
library(sessioninfo)
library(rlang)
rlang::global_handle()
library(lobstr)
cli::pretty_print_code()
```

# LLM Disclosure

The following describes the use of large language models (LLMs) in the preparation of this document:

- LLMs were used to assist with code drafting, debugging, and editing prose.
- All LLM-generated content was reviewed and verified by the author(s).
- Final responsibility for the accuracy and integrity of this work rests with the author(s).

# Profile Visualization

Here, I use `profvis` to see what might be the bottlenecks in the current git version of the code. 

## Repository Info

```{r}
#| label: git-info
git_hash <- system("git rev-parse HEAD", intern = TRUE)
git_short <- system("git rev-parse --short HEAD", intern = TRUE)
git_branch <- system("git rev-parse --abbrev-ref HEAD", intern = TRUE)
git_msg <- system("git log -1 --format=%s", intern = TRUE)
```

This document was compiled from commit **`r git_short`** on branch **`r git_branch`** (`r git_msg`).
Full SHA: `r git_hash`

## Profile Visualization

```{r}
source(here::here("dev/igsca/benchmarking_R/igsca_benchmark.R"))
```

```{r}
bigN_smallp()
```

### Big N Small p

```{r}
pv_bigN_smallp <- profvis({
  bigN_smallp()
})
```


```{r, eval = FALSE}
# TODO: Consider using this
summaryRprof()
```

### Small N Big p


### Small N Small p

### Big N Big p

#  List of potential fixes

- Differentiate argument name from parameter itself, potentially triggering look-ups that might slow things down.

# Benchmarking

As long as I'm not comparing between installed vs locally-loaded code packages, this should be legitimate.

Here, I benchmark different approaches for handling specific  bottlenecks.

```{r}

```

# Computational Environment


```{r}
#| label: session-info
sessioninfo::session_info()
```

## Packages Used

```{r}
#| label: pkg-citations
grateful::cite_packages(output = "paragraph", out.dir = ".")
```


