---
title: "IGSCA Benchmarking"
author: ""
date: today
date-format: long
format:
  html:
    theme: flatly
    toc: true
    toc-depth: 3
    toc-expand: true
    toc-location: left
    code-fold: show
    code-tools:
      source: true
      toggle: true
    code-download: true
    embed-resources: true
    max-width: 1600px
  typst:
    toc: true
bibliography: grateful-refs.bib
execute:
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: true

library(RhpcBLASctl)
RhpcBLASctl::blas_set_num_threads(2)
library(profvis)
library(proftools)
library(bench)
library(grateful)
library(sessioninfo)
library(rlang)
rlang::global_handle()
library(lobstr)
cli::pretty_print_code()
library(htmlwidgets)
library(graph)
library(Rgraphviz)
```

# LLM Disclosure

The following describes the use of large language models (LLMs) in the preparation of this document:

- LLMs were used to assist with code drafting, debugging, and editing prose.
- All LLM-generated content was reviewed and verified by the author(s).
- Final responsibility for the accuracy and integrity of this work rests with the author(s).

# Profile Visualization

Here, I use `profvis` to see what might be the bottlenecks in the current git version of the code. 

## Repository Info

```{r}
#| label: git-info
git_hash <- system("git rev-parse HEAD", intern = TRUE)
git_short <- system("git rev-parse --short HEAD", intern = TRUE)
git_branch <- system("git rev-parse --abbrev-ref HEAD", intern = TRUE)
git_msg <- system("git log -1 --format=%s", intern = TRUE)
```

This document was compiled from commit **`r git_short`** on branch **`r git_branch`** (`r git_msg`).
Full SHA: `r git_hash`

```{r}
#| label: profvis-dir
profvis_dir <- here::here(
  "dev/igsca/benchmarking_R",
  paste0(git_branch, "_", format(Sys.Date(), "%Y-%m-%d"))
)
dir.create(profvis_dir, recursive = TRUE, showWarnings = FALSE)
```

## Profile Visualization

```{r}
source(here::here("dev/igsca/benchmarking_R/igsca_benchmark.R"))
```

### Big N Small p

#### proftools

```{r}
pt_bigN_smallp <- proftools::profileExpr(bigN_smallp())
```

```{r,  eval = FALSE}
funSummary(pt_bigN_smallp) |> View()
```

```{r, eval  = FALSE}
callSummary(pt_bigN_smallp) |> View()
```

```{r,  eval = FALSE}
srcSummary(pt_bigN_smallp)
```

```{r,  eval = FALSE}
hotPaths(pt_bigN_smallp)
```

```{r}
plotProfileCallGraph(pt_bigN_smallp)
```

```{r}
flameGraph(pt_bigN_smallp)
```

```{r}
calleeTreeMap(pt_bigN_smallp)
```

#### Profvis 

```{r}
#| output: false
pv_bigN_smallp <- profvis(
  {
    bigN_smallp()
  },
  prof_output = file.path(profvis_dir, "pv_bigN_smallp.Rprof"),
  width = "100%"
)
```

```{r}
pv_bigN_smallp
```


### Small N Big p
#### proftools

```{r}
pt_smallN_smallp <- proftools::profileExpr(smallN_smallp())
```

#### Profvis 

```{r}
#| output: false
pv_smallN_smallp <- profvis(
  {
    smallN_smallp()
  },
  prof_output = file.path(profvis_dir, "pv_smallN_smallp.Rprof"),
  width = "100%"
)
```

```{r}
pv_smallN_smallp
```




### Small N Small p
#### proftools

```{r}
pt_smallN_bigp <- proftools::profileExpr(smallN_bigp())
```

#### Profvis 

```{r}
#| output: false
pv_smallN_bigp <- profvis(
  {
    smallN_bigp()
  },
  prof_output = file.path(profvis_dir, "pv_smallN_bigp.Rprof"),
  width = "100%"
)
```

```{r}
pv_smallN_bigp
```




### Big N Big p
#### proftools

```{r}
pt_bigN_bigp <- proftools::profileExpr(bigN_bigp())
```

```{r}
plotProfileCallGraph(pt_bigN_bigp)
```

```{r}
flameGraph(pt_bigN_bigp)
```

```{r}
calleeTreeMap(pt_bigN_bigp)
```

#### Profvis 

```{r}
#| output: false
pv_bigN_bigp <- profvis(
  {
    bigN_bigp()
  },
  prof_output = file.path(profvis_dir, "pv_bigN_bigp.Rprof"),
  width = "100%"
)
```

```{r}
pv_bigN_bigp
```

### Save profvis outputs

```{r}
#| label: save-profvis
profvis_outputs <- list(
  pv_bigN_smallp  = pv_bigN_smallp,
  pv_smallN_smallp = pv_smallN_smallp,
  pv_smallN_bigp  = pv_smallN_bigp,
  pv_bigN_bigp    = pv_bigN_bigp
)

for (nm in names(profvis_outputs)) {
  htmlwidgets::saveWidget(
    profvis_outputs[[nm]],
    file = file.path(profvis_dir, paste0(nm, ".html")),
    selfcontained = TRUE
  )
}
```

#  List of potential fixes

- Differentiate argument name from parameter itself, potentially triggering look-ups that might slow things down.


- Based on  both the profvis and the interpretation of profvis using Claude, the following functions should be minimized 

```{r, eval = FALSE}
%*% # A lot of time
svd() # Some of the most time
qr() # Some of the most time

ginv() # some time
solve() # some time

norm() # very little time
t()
```

## Potential Action Plan

1. Re-organize functions used for updating weights -- only two. One for updating weights of composites and another for updating weights of common factors. Also fix the assignment of the construct scores as that is computationally costly... [R]
2. Re-organize `updateCBDU()` [R]
3. Make the update of unique scores and unique loadings into a single common function across igsca and GSCA-M [R]
4. Minimize the number of transposes `t()`, `%*%` and `diag` if possible. If there's a large number of matrix multiplications to do in a zone, then consider putting  it all into a single C++ program. Also see if  the use of `MASS::ginv()` can be reduced, as it uses `svd()` and is also computationally expensive. Also, try avoiding memory intensive things where we are writing to a subset of an object. [R]
5. Improve numerical stability of updating of  loadings/path-coefficients [R]
6. Improve numerical stability of updating of weights if possible [R], or else write into C++ [C++]
7. Write update of unique/scores unique-loadings in C++ [C++]
8. Improve numerical stability for how unique scores/unique-loadings is done
9. Reduce the number of random assignments + temporary variables, if possible
10. Be careful about how RcppArmadillo and the package are using BLAS. The wasteful/slow multithreading could be a problem.

# Benchmarking

As long as I'm not comparing between installed vs locally-loaded code packages, this should be legitimate.

Here, I benchmark different approaches for handling specific  bottlenecks. It takes a long time to run, but on my home PC, it seems like 2 threads is optimal, but 1 thread is very close. On the HPC, they disable hyperthreading anyways...

## bench

```{r}
library(bench)
library(ggplot2)
library(ggbeeswarm)
```

### Correlation and A^T A of Normalized DAta

```{r}
dat <- cSEM.DGP::generateData(dgp_bigp, .empirical = TRUE, .N = 10000) |> 
  as.matrix()

# Normalize data
normed_dat <- dat/ sqrt(nrow(dat) - 1)

bm_cor_normed  <- bench::mark(
  cor(normed_dat),
  t(normed_dat) %*% normed_dat
)

```

```{r}
autoplot(bm_cor_normed)
```

### Unique Error Extraction

```{r}
RhpcBLASctl::blas_set_num_threads(2)
igsca_big <- bigN_bigp()

Z <- igsca_big$Information$Data
Z_normed <- Z / sqrt(nrow(Z) - 1)

D <- diag(ncol(Z_normed))

Eta <- igsca_big$Estimates$Construct_scores
Eta_normed <- Eta / sqrt(nrow(Eta) - 1)
n_indicators <- ncol(Z_normed)
n_case <- nrow(Z_normed)
n_constructs <- ncol(Eta)
```

```{r}
# This assumes that the expressions are equal -- they're not.
svd_comp <-  bench::mark(
  SVDA =  {
    # Path SVD A
    svd_etaprod <- svd(t(Eta_normed) %*% Eta_normed)
    gd2 <- diag(svd_etaprod$d)
    gv <- svd_etaprod$v
    GU <- Eta_normed %*% gv %*% solve(sqrt(gd2))
    M3 <- Z_normed %*% D - GU %*% (t(GU) %*% Z_normed) %*% t(D)
    svd_M3prod <- svd(t(M3) %*% M3)
    d2 <- diag(svd_M3prod$d)
    v <- svd_M3prod$v
    u <- M3 %*% v %*% solve(sqrt(d2))
    U
    u[, 1:n_indicators, drop = FALSE] %*% t(v)
  },
  SVDB = {
    # Path SVD B
    svd_etaprod <- svd(t(Eta_normed) %*% Eta_normed)
    gd2 <- diag(svd_etaprod$d)
    gv <- svd_etaprod$v
    GU <- Eta_normed %*% gv %*% solve(sqrt(gd2))
    M3 <- Z_normed %*% D - GU %*% (t(GU) %*% Z_normed) %*% t(D)
    svd_M3prod <- svd(t(M3) %*% M3)
    d2 <- diag(svd_M3prod$d)
    v <- svd_M3prod$v
    u <- M3 %*% v %*% solve(sqrt(d2))
    U
    u[, 1:n_indicators, drop = FALSE] %*% t(v)
  },
  QR_SVD = {
    Eta_Q2 <- qr.Q(qr(Eta_normed), complete = TRUE)[,
      (n_constructs + 1):n_case,
      drop = FALSE
    ]
    mx <- D %*% t(Z_normed) %*% Eta_Q2
    svd_mx <- svd(mx)
    Utilde <- svd_mx$v %*% t(svd_mx$u)
    Eta_Q2 %*% Utilde
  },
  check = FALSE
)

svd_comp
autoplot(svd_comp)
```


```{r}
# TODO: The LLM and also the SVDA/SVDB approaches diverge when there are single indicator common factors. The current igscaPrimer test fails because the loading for the customerSatisfaction is not equal to 1, which seems to make the downstream path coefficients change.
LLM_SVD_comp <- bench::mark(
  LLM = {
    # Path SVD B
    Q_thin <- qr.Q(qr(Eta_normed))
    ZD <- Z_normed %*% D
    M_proj <- ZD - Q_thin %*% crossprod(Q_thin, ZD)
    svd_mx <- svd(M_proj)
    svd_mx$u %*% t(svd_mx$v)
  },
  QR_SVD = {
    Eta_Q2 <- qr.Q(qr(Eta_normed), complete = TRUE)[,
      (n_constructs + 1):n_case,
      drop = FALSE
    ]
    mx <- D %*% t(Z_normed) %*% Eta_Q2
    svd_mx <- svd(mx)
    Utilde <- svd_mx$v %*% t(svd_mx$u)
    Eta_Q2 %*% Utilde
  },
  check = FALSE # Fails if true
)

LLM_SVD_comp
autoplot(LLM_SVD_comp)
```

#### Hidden
```{r, eval = FALSE}
# Path SVD A
svd_etaprod <- svd(t(Eta_normed) %*% Eta_normed)
gd2 <- diag(svd_etaprod$d)
gv <- svd_etaprod$v
GU <- Eta_normed %*% gv %*% solve(sqrt(gd2))
M3 <- Z_normed %*% D - GU %*% (t(GU) %*% Z_normed) %*% t(D)
svd_M3prod <- svd(t(M3) %*% M3)
d2 <- diag(svd_M3prod$d)
v <- svd_M3prod$v
u <- M3 %*% v %*% solve(sqrt(d2))
SVDA_U <- u[, 1:n_indicators, drop = FALSE] %*% t(v)


# Path SVD B
svd_etaprod <- svd(t(Eta_normed) %*% Eta_normed)
gd2 <- diag(svd_etaprod$d)
gv <- svd_etaprod$v
GU <- Eta_normed %*% gv %*% solve(sqrt(gd2))
M3 <- Z_normed %*% D - GU %*% (t(GU) %*% Z_normed) %*% t(D)
svd_M3 = svd(M3)
u <- svd_M3$u
v <- svd_M3$v
# U
u[, 1:n_indicators, drop = FALSE] %*% t(v)

# Path QR -> SVD
Eta_Q2 <- qr.Q(qr(Eta_normed), complete = TRUE)[,
  (n_constructs + 1):n_case,
  drop = FALSE
]
svd_mx <- svd(D %*% t(Z_normed) %*% Eta_Q2)
Utilde <- svd_mx$v %*% t(svd_mx$u)
Eta_Q2 %*% Utilde

# LLM
Q_thin <- qr.Q(qr(Eta_normed))
ZD <- Z_normed %*% D
M_proj <- ZD - Q_thin %*% crossprod(Q_thin, ZD)
svd_mx <- svd(M_proj)
svd_mx$u %*% t(svd_mx$v)
```

```{r}
waldo::compare(SVDA_U, QR_SVDA_U)
```

## BLAS Threads

```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(24)
bench::system_time({bigN_bigp()})
```

```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(4)
bench::system_time({bigN_bigp()})
```

```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(2)
bench::system_time({bigN_bigp()})
```

```{r, eval = FALSE}
RhpcBLASctl::blas_set_num_threads(1)
bench::system_time({bigN_bigp()})
```

# Computational Environment


```{r}
#| label: session-info
sessioninfo::session_info(info = "all")
```

## Packages Used

```{r}
#| label: pkg-citations
grateful::cite_packages(output = "paragraph", out.dir = ".")
```


