---
title: "Postestimation: Testing a model"
subtitle: "Using the `test_*()` family of functions"
date: "Last edited: `r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
vignette: >
  %\VignetteIndexEntry{using-test}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: ../inst/REFERENCES.bib
---
WARNING: this document is work in progress.

# Introduction {#introduction}

The family of test function includes `testOMF` (test for overall model fit), `testMGD`
(test for multigroup differences), `testHausman2` (Hausman test for endogeneity), and 
`testMICOM` (test for composite measurement invariance).

# Test for overall model fit (`testOMF`)
The test for overall model fit is used to test whether the specified model fits the
dataset. A bootstrap-based test suggested by is employed.

## Syntax & Options {#syntax}

```{r eval=FALSE}
testOMF(
  .object                = args_default()$.object, 
  .alpha                 = args_default()$.alpha, 
  .handle_inadmissibles  = args_default()$.handle_inadmissibles, 
  .R                     = args_default()$.R, 
  .saturated             = args_default()$.saturated,
  .seed                  = args_default()$.seed,
  .verbose               = args_default()$.verbose
  )
```

`.object`

:   An object of class `cSEMResults` resulting from a call to `csem()`.
   
`.alpha`

:  	An integer or a numeric vector of significance levels. Defaults to `0.05`.

  
`.handle_inadmissibles`

:   Character string. How should inadmissible results be treated during 
    bootstrap/permutation? One of "drop", "ignore", or "replace". 
    If "drop", inadmissibles estimations are dropped (i.e., the number of 
    results returned will be likely less than .R). For "ignore" inadmissible estimations are 
    returned even if they are inadmissible (i.e., number of results
    returned = .R). For "replace" inadmissible estimations are replaced, i.e.,
    resampling continues until there are exactly .R admissible solutions 
    (i.e., number of results returned equal .R). Defaults to "drop".

`.R`

:   Integer. The number of bootstrap runs. Defaults to `499`.


`.saturated`

:   Logical. Whether the model-implied indicator variance-covariance matrix 
    should be based on the a saturated structural model. Defaults to `FALSE` 

`.seed`

:   Integer or `NULL`. The random seed to use. Defaults to `NULL` in which case
    an arbitrary seed is choosen.

    
`.verbose`

:

# Test for multi-group differences (`testMGD`)
The test for multigroup differences is generally used to compare groups. In doing so, various 
suggested test are implemented:
- Approach suggested by @Chin1999 to compare a single parameter between two groups 

## Syntax & Options {#syntax}
```{r eval=FALSE}
testMGD(
  .object                = args_default()$.object,
  .alpha                 = args_default()$.alpha,
  .approach_p_adjust     = args_default()$.approach_p_adjust,
  .approach_mgd          = args_default()$.approach_mgd,
  .model                 = args_default()$.model,
  .handle_inadmissibles  = args_default()$.handle_inadmissibles,
  .R_permutation         = args_default()$.R_permutation,
  .R_bootstrap           = args_default()$.R_bootstrap,
  .saturated             = args_default()$.saturated,
  .seed                  = args_default()$.seed,
  .type_vcv              = args_default()$.type_vcv,
  .verbose               = args_default()$.verbose
  ) 
```

`.object`

:   An object of class `cSEMResults_multi` resulting from a call to `csem()`.
   
`.alpha`

:  	An integer or a numeric vector of significance levels. Defaults to `0.05`.

`.approach_p_adjust`

:  	Character string. Approach used for the correction of the p-value in case of 
    multiple comparisons. One of "holm", "hochberg", "hommel", "bonferroni", "BH", "BY",
   "fdr", and "none". Defaults to "none".
  
`.approach_mgd`

:  	Character string. Approach used for multi-group comparison. One of "Chin", "Sarstedt",
    and "Klesel". Defaults to "none".

`.model`

:   A model in lavaan model syntax or `NULL`. The specified model determines which parameters 
    are compared across groups. If set to `NULL` all model parameters are comapred across groups.
    Defaults to `NULL`. 

`.handle_inadmissibles`

:   Character string. How should inadmissible results be treated during 
    bootstrap/permutation? One of "drop", "ignore", or "replace". 
    If "drop", inadmissibles estimations are dropped (i.e., the number of 
    results returned will be likely less than .R). For "ignore" inadmissible estimations are 
    returned even if they are inadmissible (i.e., number of results
    returned = .R). For "replace" inadmissible estimations are replaced, i.e.,
    resampling continues until there are exactly .R admissible solutions 
    (i.e., number of results returned equal .R). Defaults to "drop".

`.R_permutation`

:   Integer. The number of permutation runs. Defaults to `499`.

`.R_bootstrap`

:   Integer. The number of bootstrap runs. Defaults to `499`.

`.saturated`

:   Logical. Whether the model-implied indicator variance-covariance matrix 
    should be based on a saturated structural model. Defaults to `FALSE` 

`.seed`

:   Integer or `NULL`. The random seed to use. Defaults to `NULL` in which case
    an arbitrary seed is choosen.

`.type_vcv`

:   Character string. Indicates which model-implied correlation matrix is compared
    across groups. One of "*indicator*" or "*construct*". Defaults to "*indicator*".
    
`.verbose`

:

# Hausman test for endogeneity (`testHausman2`)

## Syntax & Options {#syntax}
```{r eval=FALSE}
testHausman2(
 .object               = NULL,
 .alpha                = args_default()$.alpha,
 .eval_plan            = args_default()$.eval_plan,
 .handle_inadmissibles = args_default()$.handle_inadmissibles,
 .R                    = args_default()$.R,
 .resample_method      = args_default()$.resample_method,
 .seed                 = args_default()$.seed
)
```


`.object`

:   An object of class `cSEMResults` resulting from a call to `csem()`.
   
`.alpha`

:  	An integer or a numeric vector of significance levels. Defaults to `0.05`.


`.eval_plan`

:   Character string. The evaluation plan to use. One of "sequential" or 
    "multiprocess". In the latter case all available cores will be used. 
Defaults to "sequential". See `resamplecSEMResults()` for details.
   
`.handle_inadmissibles`

:   Character string. How should inadmissible results be treated during 
    bootstrap/permutation? One of "drop", "ignore", or "replace". 
    If "drop", inadmissibles estimations are dropped (i.e., the number of 
    results returned will be likely less than .R). For "ignore" inadmissible estimations are 
    returned even if they are inadmissible (i.e., number of results
    returned = .R). For "replace" inadmissible estimations are replaced, i.e.,
    resampling continues until there are exactly .R admissible solutions 
    (i.e., number of results returned equal .R). Defaults to "drop".

`.R`

:   Integer. The number of bootstrap/permutation runs. Defaults to `499`.

`.resample_method`

:   Character string. The resampling method to use. One of: "bootstrap" or "jackknife". 
    Defaults to "bootstrap".


`.seed`

:   Integer or `NULL`. The random seed to use. Defaults to `NULL` in which case
    an arbitrary seed is choosen.



# Test for measurement invariance of composite models (`testMICOM`)

## Syntax & Options {#syntax}

```{r eval=FALSE}
testMICOM(
  .object               = args_default()$.object,
  .alpha                = args_default()$.alpha,
  .approach_p_adjust    = args_default()$.approach_p_adjust, 
  .handle_inadmissibles = args_default()$.handle_inadmissibles,
  .R                    = args_default()$.R,
  .seed                 = args_default()$.seed,
  .verbose              = args_default()$.verbose
  )
```

`.object`

:   An object of class `cSEMResults` resulting from a call to `csem()`.
   
`.alpha`

:  	An integer or a numeric vector of significance levels. Defaults to `0.05`.

`.approach_p_adjust`

:  	Character string. Approach used for the correction of the p-value in case of 
    multiple comparisons. One of "holm", "hochberg", "hommel", "bonferroni", "BH", "BY",
   "fdr", and "none". Defaults to "none".

`.handle_inadmissibles`

:   Character string. How should inadmissible results be treated during 
    bootstrap/permutation? One of "drop", "ignore", or "replace". 
    If "drop", inadmissibles estimations are dropped (i.e., the number of 
    results returned will be likely less than .R). For "ignore" inadmissible estimations are 
    returned even if they are inadmissible (i.e., number of results
    returned = .R). For "replace" inadmissible estimations are replaced, i.e.,
    resampling continues until there are exactly .R admissible solutions 
    (i.e., number of results returned equal .R). Defaults to "drop".

`.R`

:   Integer. The number of bootstrap/permutation runs. Defaults to `499`.


`.seed`

:   Integer or `NULL`. The random seed to use. Defaults to `NULL` in which case
    an arbitrary seed is choosen.


`.verbose`

:   
## Details

Details if there are any.

# Usage & Examples {#examples}

## Example 1 {#example1}
## Example 2 {#example2}
## Example 3 {#example3}

# Methods & Formulae {#methods}

## Test for overall model fit {#topic1}
### Definition

### Formula

### Implementation

### See also

## Test for multigroup differences {#topic2}
### Definition

### Formula

### Implementation

### See also

## Hausman test for endogeneity {#topic3}
### Definition

### Formula

### Implementation

### See also

## Test for composite measurement invariance {#topic4}
### Definition

### Formula

### Implementation

### See also

# Literature
